---
title: Report on study from the `AusTraits` data compilation
output:
  html_document:
    df_print: kable
    highlight: tango
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

<!-- hack to get indentation on 3rd level of floating TOC; see
https://stackoverflow.com/questions/46201753/rmarkdown-indentation-of-toc-items-in-html-output
 -->

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# knitr defaults
root.dir = rprojroot::find_root("remake.yml")
knitr::opts_knit$set(root.dir = root.dir)
knitr::opts_chunk$set(echo=FALSE, cache=FALSE)

# default for table format
options(knitr.table.format = "html")

# remove warnings from dplyr about "summarise"
options(dplyr.summarise.inform = FALSE)

# Guidelines for writing report code
# - use tidyverse style and format: http://htmlpreview.github.io/?https://github.com/nicercode/2018_BEES_regression/blob/master/tidyverse.html
# - use kableExtra for styling: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# - use knitr chunck options: https://rmarkdown.rstudio.com/lesson-3.html

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
library(austraits.build)
library(austraits)
library(knitr)
library(kableExtra)

my_kable_styling <- my_kable_styling_html 

## Assumes to items exist in global name space
if(!exists("austraits")) {
  austraits <- remake::make("austraits")
  # stop("austraits must exist in global name space to knit a report")
}
if(!exists("dataset_id")) {
  dataset_id <- "Westoby_2014"
  #  stop("dataset_id must exist in global name space to knit a report")
}

trait_definitions <- austraits$trait_definitions
data_study <- extract_dataset(austraits, dataset_id)
metadata <- read_metadata(sprintf("data/%s/metadata.yml", dataset_id))
schema <- austraits$schema

knitr::opts_chunk$set(fig.cap='', fig.path = file.path(root.dir,"export/figures", paste0(dataset_id,"_")))

# start notetaker device
questions <-  austraits.build:::start_notetaker()

new_question <- function(txt) {
  x <- questions( austraits.build:::add_note,  austraits.build:::as_note(txt)) %>%  austraits.build:::print_notes(as_anchor=TRUE)
  writeLines(sprintf("<font color='orange'>**Question:** %s</font>\n\n", paste0(x, collape="")))
}

# Variable to record when definitions have been specified

structures_already_defined <- c("names")

trait_type <- function(trait_name, trait_definitions) {
 extract_list_element(trait_name, trait_definitions, "type")
}

trait_is_numeric <- function(trait_name, trait_definitions) {
 trait_type(trait_name, trait_definitions) == "numeric"
}

trait_is_categorical <- function(trait_name, trait_definitions) {
 !trait_is_numeric(trait_name, trait_definitions)
}

```

# Introduction

## About AusTraits

`AusTraits` is an open-source compilation of trait data for Australian plant species. The traits we are considering are any morphological, functional or physiological variable that: 1. Tend to vary widely among species; and 2. Have been collected across a moderate number of species. Many people collect trait data, and as such, data is spread across many sources, under a variety of formats and terms.

AusTraits combines all of these data fragments into a harmonised, error checked, publicly available compilation.

AusTraits is built from many different sources, here referred to as `studies` (so called because the sources are most-often individual scientific papers), with each study denoted by a distinct `dataset_id`. Our data processing pipeline seeks to combine all of the different studies in a transparent and reproducible way. In addition, the AusTraits repository consists of configuration files defining the list of known species, the definitions for each trait and table, and appropriate unit conversions. Detailed descriptions of the `trait_definitions`, `database structure`, and `file format` are all available on the [project homepage](http://traitecoevo.github.io/austraits.build/) under the `Details` tab.

## Purpose of this report

The AusTraits compilation includes data from a study with dataset_id **` `r dataset_id` `**. This document describes the information we have on that study.
This report gives both you, as a data contributor, and us, as AusTraits compilers, the opportunity to:

1. Verify that our understanding and handling of your data is correct
2. Provide any missing information about the data

Data from each study in AusTraits is organised into two files, within its own folder:

- ` data/`r dataset_id`/data.csv`: contains the primary trait measurements, in either long or wide format
- ` data/`r dataset_id`/metadata.yml`: contains all the contextual data (or metadata) about the trait measurements.

Data from those files has been used to generate the report below.

Please review the report and the original files, check they are correct, and send updates where appropriate. Please also answer the queries in the report below, appearing in orange. These appear again as a list at the end of the document.

We would welcome any of the following:

- Answers to the list of questions
- Expanding or correcting the information in the `metadata.yml`
- Updates to the data

To aid your review, we have attached the following files:

- `r dataset_id`/data.csv` and ` `r dataset_id`/metadata.yml`: The raw inputs for your study.

### Send feedback by 

1. **Copying and pasting the questions at the end of the report into an email and answering as many as possible.** 

    The majority of the questions are automatically generated prompts to look at the pertinent metadata or data plot (or table) and confirm it matches your expectations - for these simply answer "yes". There are often a few questions targeted to your study, written by an AusTraits data curator as they merged in your study. Please consider these carefully, as they usually indicate missing metadata or some source of ambiguity (e.g. units not provided, uncertainty about trait value substitutions).    
2. If appropriate, send a new spreadsheet to replace the existing `data.csv` file.
3. You may edit the `metadata.yml` file associated with your study yourself, but most contributors prefer to answer the flagged questions and have an AusTraits data curator make the changes to the file.
4. Sending your responses and any edited files to our email [austraits.database@gmail.com](mailto:austraits.database@gmail.com).


```{r, results='asis', echo=FALSE}
# Any additional questions from metadata

if(!is.null(metadata$questions) & !is.na(metadata$questions)) {

  x <- names(metadata$questions)[ !(names(metadata$questions) %in% c(
                                      "additional_traits", "austraits") 
                                ) ]
  
  if(length(x) > 0)                            
    writeLines("# Major questions\n\n")

  for(i in seq_along(x))
    new_question(sprintf("(section `general`) %s", metadata$questions[[i]]))
}
```


# Study details

We collect information about the context of the trait measurements. This includes details on the people involved, the sources, and details about type of collection, including location and time of collections.

## People

The following people are listed as involved in this study:

```{r, results='asis', echo=FALSE}
metadata$contributors$data_collectors %>%
  list_to_df() %>%
  my_kable_styling()
```

If a current affiliation isn't known, provide the last known institution. We also require an email address for all people. People involved in the study who cannot be tracked down, need to be designated as assistants.

`r `metadata$contributors$austraits_curators` entered the data into AusTraits.

```{r, results='asis', echo=FALSE}
new_question("(section `people`) Are all appropriate people listed, with appropriate details?")
```

## Sources

Data from this study is recorded as coming from the following sources:

Primary citation:

```{r, results='asis', echo=FALSE}
metadata$source$primary %>%
  list1_to_df() %>%
  my_kable_styling()
```

```{r, results='asis', echo=FALSE}
n <- length(metadata$source)
if(n > 1) {
  writeLines("Secondary citations:")

  for(i in 2:n) {
    metadata$source[[i]] %>%
    list1_to_df() %>%
    my_kable_styling() %>% 
    cat()
    writeLines("\n\n")
  }
}
```


```{r, results='asis', echo=FALSE}
new_question("(section `source`) Are the citation details for this study correct?")
```

## General description

The description we have of the study is as follows:

```{r, results='asis', echo=FALSE}
metadata$dataset %>%
  list1_to_df() %>%
  my_kable_styling()
```

```{r, results='asis', echo=FALSE}
new_question("(section `dataset`) Can you provide more detailed information for any of these variables?")
```

```{r, results='asis', echo=FALSE}
missing <- metadata$dataset %>%
 list1_to_df() %>%
 filter(value == "unknown") %>%
 pull(key)

for(v in missing) {
  sprintf("(section `dataset`) Can you provide missing details for the variable `%s`?\n", v) %>%
    new_question()
  sprintf("The variable `%s` captures *'%s'*\n", v, definitions$metadata$elements$dataset$values[[v]]) %>%
    writeLines()
}
```


## Sites

Data were collected at the following sites and locations. As a minimum, we are aiming to collect, the variables

```{r, results='asis', echo=FALSE, warning=FALSE}

create_if_missing <- function(df, var, type=rep(NA_character_, nrow(df))) {
  if(is.null(df[[var]]))
    df[[var]] <- type
  df
}

site_main <- c("longitude (deg)","latitude (deg)", "description")
sites <-
  full_join(by="site_name",
            # count of records by site
            data_study$traits %>%
              mutate(site_name = ifelse(is.na(site_name), "unknown", site_name)) %>%
              group_by(site_name) %>%
              summarise(records = n()),
            # details on sites from metadata
            data_study$sites %>%
              filter(site_property %in% site_main) %>%
              select(-dataset_id) %>%
              spread(site_property, value) %>%
              create_if_missing("site_name") %>%
              create_if_missing("latitude (deg)") %>%
              create_if_missing("longitude (deg)") %>%
              create_if_missing("description")  %>%
              mutate(
                `latitude (deg)` = as.numeric(`latitude (deg)`),
                `longitude (deg)` = as.numeric(`longitude (deg)`)
                )
  ) %>%
 select(site_name, records, `latitude (deg)`, `longitude (deg)`, description)

 sites %>%
   my_kable_styling()
         

if( any(tolower(sites$site_name) %in% c(NA, "unknown") ) ) {
  new_question("(section `sites`) **Location data incomplete or unknown!** Can you provide location details where your data were sampled? If data are from a National Park, biological reserve, or other location with a name, but not linked to a specific site within the park, the site name should be that national park (or similar), and we will use the park's headquarters as the GPS coordinates. If the data are from herbarium specimens or a species monograph it is appropriate to leave this information blank.")
}

```

```{r maps, eval= TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.keep="all"}

# Make a map of study sites using leaflet
sites <- sites %>% filter(!is.na(`latitude (deg)`) & !is.na(`longitude (deg)`))
if(nrow(sites) > 0 ) {
  leaflet::leaflet() %>%
  leaflet::addTiles() %>%
  leaflet::addScaleBar() %>%
  leaflet::addMiniMap(zoomLevelFixed=2) %>%
  leaflet::addMarkers(lng = sites$`longitude (deg)`, lat = sites$`latitude (deg)`,
             label = sites$site_name)	
}
```

```{r, results='asis', echo=FALSE}
new_question("(section `sites`) Do site details look complete and accurate? As a minimum we would like `latitude`, `longitude`, `description`.")
```

## Additional site data (optional)

The following variables have been recorded for the sites included in the study:

```{r, results='asis', echo=FALSE}

data_study$sites %>%
  filter(!(site_property %in% site_main)) %>%
  select(-dataset_id) %>%
  spread(site_property, value) -> sites_extra

if(nrow(sites_extra) > 0 ) {
  sites_extra %>%
    my_kable_styling() %>%
    writeLines()

  new_question("(section `sites`) Do the additional site details look complete?")
  } else{
  writeLines("There is currently no additional site data associated with this dataset.")
}

```

## Contexts


The following contexts have been recorded for the measurements included in this study. The `type` column has been added to easily filter records across many studies.

```{r, results='asis', echo=FALSE}
data_study$contexts %>%
  select(-dataset_id) %>%
  spread(context_property, value) -> context

if(nrow(context) > 0 ) {
  context %>%
    # make ordering of columns, same as dataframe
    select(one_of(c("context_name", data_study$contexts$context_property %>% unique()))) %>%
    my_kable_styling() %>%
    writeLines()

  new_question("(section `contexts`) Do the context details look complete?")
  } else{
  writeLines("There is currently no contextual data associated with this dataset.")
}

```

## Species sampled

We have records on `r data_study$traits$taxon_name %>% unique() %>% length()` species from your study. We have attempted to align species names with [the Australian Plant Census  (APC)](https://biodiversity.org.au/nsl/services/APC) -- a global working list of all known Australian vascular plant species. The full list of names for species recorded in your study are listed below.

# Trait measurements

## Overview

The dataset includes `r data_study$traits %>% nrow()` individual data points from `r data_study$traits$taxon_name %>% unique() %>% length()` species, with data included for the following `r data_study$traits$trait_name %>% unique() %>% length()` traits:

```{r summary_table, results='asis', echo=FALSE}
n_records <- data_study$traits %>%
  group_by(trait_name) %>%
    summarise(
        species = length(unique(taxon_name)),
        records = length(value)) %>%
   mutate(records_per_species = round(records / species, digits=2) )

n_records %>%
  my_kable_styling()
```


```{r, results='asis', echo=FALSE}
new_question("(section `traits`) Does this study include other trait data we may have missed?")

if(median(n_records$records_per_species) < 2) {
  new_question("(section `traits`) It appears as though we mostly have means for each species. Can you provide individual-level measurements?")
}

new_question("(section `traits`) Were any of your data sourced from other studies? If so, can you tell us which records and the source (so that we can avoid duplicates, where possible)?")

```

```{r excluded,  results='asis', echo=FALSE, warning = FALSE}

writeLines(ifelse(nrow(subset(data_study$excluded_data, error != "Missing value")) == 0,
        "All data passed quality control so nothing was excluded in this study. (Yay!)",
        "We excluded some records in your dataset as they did not pass the quality control. A summary of the number of points and reasons for exclusion are as follows:"))

if(nrow(subset(data_study$excluded_data,error!="Missing value")) > 0 ) {
  data_study$excluded_data %>%
    filter(error!="Missing value") %>%
    group_by(trait_name, error) %>%
    summarise(`points excluded` = n()) %>%
    ungroup() %>%
    my_kable_styling() %>%
    writeLines()

  writeLines(ifelse(nrow(subset(data_study$excluded_data, error == "Unsupported trait value")) > 0,
    "In particular, the following data points were excluded because the supplied trait value could not be aligned with the allowable values in AusTraits (error = `Unsupported trait value`). The allowable values for each trait are listed in the [definitions file](http://traitecoevo.github.io/austraits.build/articles/Trait_definitions.html) and lower in this report.",""))


  if(nrow(subset(data_study$excluded_data,error == "Unsupported trait value")) > 0 ) {
    data_study$excluded_data %>%
      filter(error=="Unsupported trait value") %>%
      select(trait_name,value) %>%
      rename(`value that is unaligned` = value) %>%
      distinct() %>%
      arrange(trait_name, `value that is unaligned`) %>%
      my_kable_styling() %>%
      writeLines()
  }

  writeLines(ifelse(nrow(subset(data_study$excluded_data, error == "Observation excluded in metadata")) > 0 ,
      "The following observations were explicitly excluded (error = `Observation excluded in metadata`). This decision was most likely made by the data curators because the study included non-native, not naturalised taxa or species that aren't vascular plants.",""))
"  "

  if(nrow(subset(data_study$excluded_data,error == "Observation excluded in metadata")) > 0) { 
    metadata$exclude_observations %>%
      list_to_df() %>%
      my_kable_styling() %>%
      writeLines()
  }

  writeLines("Summaries of excluded numeric observations are included later in the report.")

  new_question("(section `traits`) Can you provide any additional information so that above exclusions (required substitutions or other errors) no longer apply?")
}
```


## Numerical traits

```{r numerical, results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10, eval=TRUE}
traits <- data_study$traits$trait_name %>% unique() %>% sort()
traits_numeric <- traits[trait_is_numeric(traits, trait_definitions)]



question_x <- function(x, y = trait) {
  sprintf("(section `traits`) Can you provide missing details `%s` for trait `%s`?\n", x, y) %>%
  new_question()
}

define_variable <- function(var, schema) {

  description <- switch(var,
                        value_type = definitions$value_type$description,
                        replicates = schema$austraits$elements$traits$elements$replicates,
                        methods = schema$austraits$elements$methods$elements$methods
                        )

  sprintf("The variable `%s` is *'%s'*.", v, description) %>%
     writeLines()
  
  if(var == "value_type") {
    writeLines("Possible values are:")

    schema$value_type$values %>%
      list1_to_df()  %>%
      my_kable_styling() %>%
      writeLines()
    }
}



writeLines(ifelse(length(traits_numeric) == 0,
        "No numerical traits data are currently available in this dataset",
        "We have recorded the following traits with continuous numerical values in this dataset. Plots are shown comparing the distribution of values in your study (blue) to those in other datasets within AusTraits (red, green). For each trait, please eyeball the information provided and let us know of any issues or updates where appropriate."))

for(trait in traits_numeric) {


  data_trait_study <- data_study$traits %>%
    filter(trait_name == trait) %>%
    mutate(value = as.numeric(value))

  data_trait_all <- austraits$traits %>%
    filter(trait_name == trait) %>%
    mutate(value = as.numeric(value))

  transforms <- metadata$traits %>% list_to_df() %>% filter(trait_name == trait)

  elements <- trait_definitions[[trait]]

  units <- elements$units

  c("", "",
    sprintf("### %s", trait),
    "",
    sprintf("We aligned the contributed variable called `%s` in the data you supplied with the trait `%s` in the AusTraits database, with value_types `%s`.\n\nThe original variable was supplied with units `%s`; these were converted to our standard `%s`.\n\nThe data has the following properties:", transforms$var_in %>% paste0(collapse=", "), trait, transforms$value_type %>% paste0(collapse=", "), transforms$unit_in[1], elements$units),
    "",
    sprintf("- **standardised name**: %s", trait ),
    sprintf("- **standardised description**: %s", elements$description ),
    sprintf("- **label**: %s", elements$label ),
    sprintf("- **units**: %s", units ),
    sprintf("- **records**: %s in this study (of %s in AusTraits)", data_trait_study %>% nrow(), data_trait_all %>% nrow() ),
    sprintf("- **allowable range**: %s - %s %s", elements$values$minimum,
            elements$values$maximum, units),
    sprintf("- **observed range in this study**: %s - %s %s", data_trait_study %>% dplyr::pull(value) %>% min(),
            data_trait_study %>% dplyr::pull(value) %>% max(), units),
    "", "",
    "Data for this trait in this study were collected using the following methods:", "",
    sprintf("- **value_type**: %s", transforms$value_type %>% paste0(collapse=", ")),
    sprintf("- **replicates**: %s", transforms$replicates %>% paste0(collapse=", ")),
    sprintf("- **methods**: %s", transforms$methods[1]),
    "","") %>%
  writeLines()

  writeLines("  ")

  writeLines(ifelse(nrow(filter(data_study$excluded_data, error == "Value out of allowable range" & trait_name == trait)) > 0,
    "The following data points were excluded because the values were outside the allowable range (error = `Value out of allowable range`). Our ranges of allowable values are sufficiently generous, that most excluded data are well outside the expected range for the trait, but if you believe these data are valid, accurate measurements that will be useful to future AusTraits users, let us know.",""))

if(nrow(filter(data_study$excluded_data, error == "Value out of allowable range" & trait_name == trait)) > 0 ) {
  data_study$excluded_data %>% filter(error == "Value out of allowable range" & trait_name == trait) %>%
    select(observation_id, taxon_name, site_name, context_name, value, unit) %>%
    my_kable_styling() %>%
    writeLines()
}


 # austraits::plot_trait_distribution_beeswarm(austraits, trait, "dataset_id", highlight=dataset_id, hide_ids = TRUE)


  writeLines(c(""))

  outliers <- data_trait_all %>%
    arrange(value) %>%
    mutate(per = seq_len(n())/n()*100) %>%
    filter(dataset_id %in% unique(data_trait_study$dataset_id),
           per < 2.5 | per > 97.5)

  # check for outliers
  if(nrow(outliers) > 0 ) {
    writeLines(sprintf("OUTLIERS: %s of points in your dataset lie in either the bottom or top 2.5 percent of data\n", nrow(outliers)))
    percent_outliers <- nrow(outliers)/nrow(data_trait_study)*100
    if(percent_outliers > 10) {
      new_question(paste0("(section `traits`) More than 10% of your data points for the trait `", trait, "` are outliers, does this seem reasonable, given what you know about the biology of these species and overall distribution of values in AusTraits?"))
    } else {
      new_question(
    sprintf("(section `traits`) Do the data for the trait `%s` appear correct?", trait)
    )
    }
  }

  # Check if methods data are complete
  methods_missing <- transforms[1,] %in% c("unknown")
  names_missing <-  names(transforms)[methods_missing]

  for(v in c("value_type", "replicates", "methods")) {

    if(v %in% names_missing ) {
      question_x(v, trait)

      if(!(v %in% structures_already_defined)) {
        define_variable(v, schema)
        structures_already_defined <- c(v, structures_already_defined)
      }
    }
  }


}
```

## Categorical traits

```{r categorical, results='asis', echo=FALSE, eval=TRUE}

traits <- data_study$traits$trait_name %>% unique() %>% sort()
traits_categorical <- traits[!trait_is_numeric(traits, trait_definitions)]

writeLines(
  ifelse(length(traits_categorical) == 0,
    "No categorical traits data are currently available in this dataset.",
    "We have recorded the following traits with categorical values in this dataset. Tables are shown comparing the distribution of values in your study to others in the dataset"))

for(trait in traits_categorical) {

  transforms <- metadata$traits %>% list_to_df() %>% filter(trait_name == trait)
  substitutions_for_trait <- metadata$substitutions %>% list_to_df() %>% filter(trait_name == trait)

  data_trait_study <- data_study$traits %>%
    filter(trait_name == trait)

  data_trait_all <- austraits$traits %>%
    filter(trait_name == trait)

  trait_definitions <- austraits$definitions

  elements <-trait_definitions$elements[[trait]]

  x <- c("", "",
    sprintf("### %s", trait),
    "",
    sprintf("We aligned the variable called `%s`in the data you supplied with the trait `%s` in the AusTraits database.\n\nThe contributed data has the following properties:", transforms$var_in, trait),
    "",
    sprintf("- **standardised name**: %s", trait ),
    sprintf("- **standardised description**: %s", trait_definitions$elements[[trait]]$description ),
    sprintf("- **label**: %s", elements$label ),
    sprintf("- **records**: %s in this study (of %s in AusTraits)", data_trait_study %>% nrow(), data_trait_all %>% nrow() ),
    "", 
    "Data for this trait in this study were collected using the following methods:", "",
    sprintf("- **value_type**: %s", transforms$value_type),
    sprintf("- **replicates**: %s", transforms$replicates),
    sprintf("- **methods**: %s", transforms$methods),
    "",
    "The allowable values for this trait in AusTraits are:",
    "",
    elements$values %>% unlist() %>% as.data.frame() %>% tibble::rownames_to_column() %>% rename(trait_value = "rowname", definition = ".") %>% arrange(trait_value) %>% my_kable_styling(),
    "",
    ifelse(nrow(substitutions_for_trait)>0,
    "The following substitutions have been implemented to align categorical trait values in the submitted data table with allowable trait values specified in the AusTraits definitions file.",
    "No substitutions were required to align the data for this trait with AusTraits allowed trait values."),
    ifelse(nrow(substitutions_for_trait)>0,
    substitutions_for_trait %>% rename("value in contributed data table"=find,"AusTraits aligned value"=replace)%>% my_kable_styling(),""),
    "The following table shows distribution of recorded values in your study and AusTraits as a whole. Note that space-delimited strings of values, indicate when multiple trait values were recorded for a single `observation` (species or individual-level) within a single study.",
    "") %>%
  writeLines()

  right_join(
    data_trait_study %>% group_by(value) %>% summarise(study = n()),
    data_trait_all %>% group_by(value) %>% summarise(total = n()),
    by= "value"
    ) %>%
  mutate(
    study = ifelse(is.na(study), 0, study),
    `% study` = round(study/sum(study)*100),
    `% total` = round(total/sum(total)*100) 
  ) %>%
  arrange(-`study`,-`total`) %>%
  my_kable_styling() %>%
  add_header_above(c(" ", "Counts" = 2, "Percent" = 2)) %>%
  column_spec(1, border_right = TRUE) %>%
  column_spec(3, border_right = TRUE) %>%
  column_spec(5, border_right = TRUE) %>%
  writeLines()

  writeLines("\n")
  new_question(
    sprintf("(section `traits`) Do the data for the trait `%s` appear correct?", trait)
  )

  # Check if methods data are complete
  methods_missing <- transforms[1,] %in% c("unknown")
  names_missing <-  names(transforms)[methods_missing]

  for(v in c("value_type", "replicates", "methods")) {

    if(v %in% names_missing ) {
      question_x(v, trait)

      if(!(v %in% structures_already_defined)) {
        define_variable(v, schema)
        structures_already_defined <- c(v, structures_already_defined)
      }
    }
  }

}
```

# Taxon list

We have records on the following taxa from your study. We have attempted to align taxon names in this study with the Australian Plant Census (APC) or Australian Plant Names Index (APNI). 

## Full taxa list

The table below shows the table of all taxa in this study, where possible using the name aligned with the APC. Columns are as follows:

```{r, results='asis', echo=FALSE}
schema$austraits$elements$taxa$elements %>%
  list1_to_df() %>%
  my_kable_styling()
```

Clicking on the name will take you to the relevant taxonomic record in the APC or APNI, where known:

```{r species_list2, results='asis', echo=FALSE}
data_study$taxa %>%
  select(one_of(names(schema$austraits$elements$taxa$elements))) -> tmp

if(nrow(tmp) < 10000 ){
  tmp %>%
  mutate(
    taxon_name = as_link(acceptedNameUsageID, taxon_name)) %>%
  select(-acceptedNameUsageID, -ccAttributionIRI) %>%
  mutate_all(replace_na, "") %>%
  my_kable_styling()
} else{
  cat(crayon:red("\t\tTable too large to dsiplay\n"))
  tmp %>%
  mutate_all(replace_na, "") %>% 
  write_csv(file.path(root.dir,"export/reports", paste0(dataset_id,"_taxa.csv")))
}
```

## Taxonomic changes

To create the list of aligned species, we needed to make some taxonomic changes. This involved two stages. 

* **Stage 1**: Where possible (i.e. there was no or only a few characters difference), the name you supplied was matched automatically with a known name in APC or APNI. In other cases we may have aligned the species by searching for an appropriate match. Such changes are documented in the study metadata file.  The variable `cleaned_name` shows the updated name. The variable `d1` shows the number of characters difference between the `original_name` and `cleaned_name`. 
* **Stage 2**: Once aligned with a known name, we used the APC to update the `cleaned_name` to an accepted name. The taxonomic status of the cleaned name is indicated in column `status cleaned name`. If accepted, no change was made. If it is a synonym or otherwise, the name was changed according to the recommendation given in the APC. Where they existed, we preferred to take the accepted status of an `cleaned_name`, if it existed. Alternative status values are indicated in brackets. This indicate if alternative uses of the name were ever applied. 

Links on `cleaned_name` and `taxon_name` take you to the APC or APNI record for that name.

```{r}
data_study$taxonomic_updates %>% select(-dataset_id) %>% 
  filter(original_name !=taxon_name) %>% 
  mutate(
    d1 = purrr::map2_dbl(original_name, cleaned_name, ~adist(.x, .y)) %>% as.character() %>% str_replace("0", ""),
    d2 = purrr::map2_dbl(taxon_name, cleaned_name, ~adist(.x, .y)) %>% as.character() %>% str_replace("0", ""),
    cleaned_name = as_link(taxonIDClean, cleaned_name),
    taxon_name = as_link(acceptedNameUsageID, taxon_name),
    alternativeTaxonomicStatusClean = replace_na(alternativeTaxonomicStatusClean, ""),
    `status cleaned name` = sprintf("%s %s", taxonomicStatusClean, ifelse(alternativeTaxonomicStatusClean=="", "", paste("(", alternativeTaxonomicStatusClean, ")")))
    ) %>% 
  select(original_name, d1, cleaned_name, d2, taxon_name, `status cleaned name`) %>%
  mutate_all(replace_na, "") %>% 
  my_kable_styling()
```

## Unknown species

We were unable to find taxonomic matches for the following species in the data. Where genus and family are shown, the specified genus aligns with an accepted genus. The term `unplaced` appears as the `status` for names on the APNI list, but not yet reviewed by the APC.

```{r species_list3, results='asis', echo=FALSE}
data_study$taxa %>%
  filter(is.na(acceptedNameUsageID)) %>%
  select(one_of(names(schema$austraits$elements$taxa$elements))) %>%
  select(-acceptedNameUsageID, -ccAttributionIRI, -taxonDistribution) %>%
  mutate_all(replace_na, "") %>%
  my_kable_styling()
```


```{r, results='asis', echo=FALSE}
new_question("(section `taxonomic_updates`) Do these taxonomic alignments and corrections look reasonable?")
```

# Review of questions

Here is a summary of the questions we would like answers to. Clicking on the links will take you back to the appropriate spot in the document. Please copy and past these into an email when you respond.

<font size = '3' color='orange'>


```{r, results='asis', echo=FALSE}
questions() %>% austraits.build:::print_all_notes() %>% writeLines()
```


</font>


# Session information

The technical information below may useful for us for the sake of reproducibility, but you don't need to worry about it!

<font color='grey'>
This report was generated using the data from the commit `r get_SHA_link()`

And the following R environment

```{r, echo=FALSE}
sessionInfo()
```

</font>

