---
title: Structure of AusTraits data compilation
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structure of AusTraits data compilation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = "asis",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(austraits.build)

my_kable_styling <- util_kable_styling_html
```

```{r, echo=FALSE, results='hide', message=FALSE}
## Loads austraits into global name space
austraits <- austraits:::austraits_lite
schema <- get_schema()
```


This document describes the structure of the AusTraits compilation, corresponding to **Version `r austraits$build_info$version`** of the dataset. 

Note that the information provided below is based on the information provided within the file `austraits.build_schema.yml`, which can be accessed by running `get_schema` or `system.file("support", "austraits.build_schema.yml", package = "austraits.build")`. 

AusTraits is essentially a series of linked components, which cross link against each other::

```{r, results="hide", comment = '', eval=FALSE}
names(austraits) %>%
  create_tree_branch("austraits.build") %>%
  writeLines()
```

```
austraits
├── traits
├── sites
├── contexts
├── methods
├── excluded_data
├── taxonomic_updates
├── taxa
├── contributors
├── sources
├── definitions
├── schema
└── build_info
```

These include all the data and contextual information submitted with each contributed dataset. 

# Components

The core components are defined as follows.

```{r}
print_schema_element <- function(elements) {
  if (elements$type == "character") {
    sprintf("**Content:** %s\n", elements$value) %>%
      writeLines()
  }

  if (elements$type == "table") {
    sprintf("**Content:** \n") %>%
      writeLines()

    elements$elements %>%
      util_list_to_df1() %>%
      my_kable_styling() %>%
      writeLines()
  }
}
```

## traits {#traits}
```{r}
elements <- schema$austraits$elements$traits

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

**Details:**  

#### Entity IDs, entity type, observation number, method number, and measurement remarks

Each trait measurement has an associated `entity_id` and `entity_type`.  Entity IDs bind together measurements within a dataset made on the same entity, and thereby allow transformation between long format (e.g. with variables `trait_name` and `value`) and wide format (e.g. with traits as columns). This is possible, because when multiple traits were collected on the same entity, they all share the same entity_id.

In detail:

- `entity_type` is `r tolower(schema$metadata$elements$traits$elements$entity_type)` Possible values are: 

```{r entity_type}
schema$entity_type$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

- `entity_id` is built from up to three segments, a taxon_id segment (for all entity types), a population_id segment (when `entity_type` is `population` or `individual`), and a individual_id segment (when `entity_type` is `individual`).

    - The taxon_id segment numbers an alphabetised list of taxa within the study sequentially, producing segments `tax001`, `tax002`, etc.
    
    - The population_id segments numbers an alphabetised list of sites or, in some cases, site by context combinations, sequentially, producing segments `pop01`, `pop02`, etc. For experimental and manipulative studies, "population" is defined as a group of individuals subjected to the same treatment conditions. 
    
    - The individual_id segments number individuals within a taxon by population combination sequentially, producing segments `ind001`, `ind002`, etc. For datasets submitted to AusTraits in long format, if a column specifying `individual_id` does not exist, it is assumed that `species_name` is the appropriate grouping variable. For datasets submitted to AusTraits in long format, if a column specifying `individual_id` does not exist, it is assumed that each row of data represents a unique entity.
    
- `observation_number` is used to indicate that multiple observations were made on the same entity for the same trait, representing repeated measures over time.

- `method_number` is used to indicate that multiple observations were made on the same entity for the same trait at a given time, using multiple methods. The difference between methods is indicated under `measurement_remarks`. In literature compilations, `method_number` can be used to distiguish between trait values for the same entity recorded from different sources.

- `measurement_remarks` is used to document `r tolower(schema$metadata$elements$traits$elements$measurement_remarks)`

#### Life stage, basis of record 

- `life_stage` is `r tolower(schema$metadata$elements$traits$elements$life_stage)`.

- `basis_of_record` is `r tolower(schema$metadata$elements$traits$elements$basis_of_record)` Possible values are:

```{r basis_of_record}
schema$basis_of_record$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

#### Values, Value types, basis of value {#value_types}

Each record in the table of trait data has an associated `value`,  `value_type`, and `basis_of_value`. 

Traits are either `numeric` or `categorical`. For traits with numerical values, the recorded value has been converted into standardised units and we have checked that the value can be converted into a number and lies within the allowable range. For categorical variables, we only include records whose values are included as allowable values (terms) in a trait's definition in the definitions table. Moreover, we use a format whereby

- we use `_` for multi-word terms, e.g. `semi_deciduous`  
- we use a space for situations where two values co-occur for the same entity. For instance, a flora might indicate that a plant species can be either annual or biennial, in which case the trait is scored as `annual biennial`.

Each trait measurement has an associated `value_type`, which is `r tolower(schema$metadata$elements$traits$elements$value_type)` Possible values are:

```{r value_type}
schema$value_type$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

Each trait measurement also has an associated `basis_of_value`, which is `r tolower(schema$metadata$elements$traits$elements$basis_of_value)` Possible values are: 

```{r basis_of_value}
schema$basis_of_value$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```


AusTraits does not include intra-individual observations made at a single point in time. When multiple measurements per individual are submitted to AusTraits, we take the mean of the values and record the value_type as `mean` and indicate under replicates the number of measurements made.  




## sites
```{r}
elements <- schema$austraits$elements$sites

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

## contexts
```{r}
elements <- schema$austraits$elements$contexts

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

## methods {#methods}
```{r}
elements <- schema$austraits$elements$methods

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## exluded_data
```{r}
elements <- schema$austraits$elements$excluded_data

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## taxa

**Description:** A table containing details on taxa that are included in the table [`traits`](#traits). We have attempted to align species names with known taxonomic units in the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc) and/or the [`Australian Plant Names Index` (APNI)](https://biodiversity.org.au/nsl/services/APNI); the sourced information is released under a CC-BY3 license. 


Version `r desc::desc_get_field("Version")` of AusTraits contains records for `r austraits$traits$taxon_name %>% unique() %>% length()` different taxa. 


```{r}
elements <- schema$austraits$elements$taxa

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## taxonomic_updates

```{r}
elements <- schema$austraits$elements$taxonomic_updates

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


Both the original and the updated taxon names are included in the [`traits`](#traits) table.


## definitions
```{r}
elements <- schema$austraits$elements$definitions

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

**Details on trait definitions:** The allowable trait names and trait values are defined in the definitions file. Each trait is 
labelled as either `numeric` or `categorical`. An example of each type is as follows. For the full list, see the [Trait definitions vignette](http://traitecoevo.github.io/austraits.build/articles/definitions.html).


```{r, traits}
for (trait in c("specific_leaf_area", "woodiness")) {
  elements <- austraits$schema$traits$elements[[trait]]

  data_trait <- austraits$traits %>% filter(trait_name == trait)

  c(sprintf("**%s**\n\n", trait), sprintf("- label: %s", elements$label), sprintf("- description: %s", elements$description), sprintf("- number of records: %s", data_trait %>% nrow()), sprintf("- number of studies: %s", data_trait %>% pull(dataset_id) %>% unique() %>% length()), sprintf("- type: %s%s", elements$type, ifelse(elements$type == "numeric", sprintf("\n- units: %s", elements$units), "")), ifelse(elements$type == "numeric", sprintf("- allowable range: %s - %s %s", elements$values$minimum, elements$values$maximum, elements$units), sprintf("- allowable values:\n%s\n", paste0("    - *", elements$values %>% names(), "*: ", elements$values %>% unlist(), collapse = "\n"))), "") %>% writeLines()
}
```


## contributors
```{r}
elements <- schema$austraits$elements$contributors

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## sources

For each dataset in the compilation there is the option to list primary and secondary citations. The primary citation is defined as, ``r austraits$schema$metadata$elements$source$values$primary$description`` The secondary citation is defined as, ``r austraits$schema$metadata$elements$source$values$secondary$description`` 

The element `sources` includes bibtex versions of all sources which can be imported into your reference library:

```
RefManageR::WriteBib(austraits$sources, "refs.bib") # write all sources to file
RefManageR::WriteBib(austraits$sources["Falster_2005_1"], "refs.bib") # write a single reference to a file
```

Or individually viewed:
```
austraits$sources["Falster_2005_1"]
```

A formatted version of the sources also exists within the table [methods](#methods). 



## build_info
```{r}
elements <- schema$austraits$elements$build_info

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```
