---
title: Structure of AusTraits data compilation
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structure of AusTraits data compilation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = "asis",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(austraits.build)

my_kable_styling <- util_kable_styling_html
```

```{r, echo=FALSE, results='hide', message=FALSE}
## Loads austraits into global name space
austraits <- austraits:::austraits_lite
schema <- get_schema()
```


This document describes the structure of the AusTraits compilation, corresponding to **Version `r austraits$build_info$version`** of the dataset. 

Note that the information provided below is based on the information provided within the file `austraits.build_schema.yml`, which can be accessed by running `get_schema` or `system.file("support", "austraits.build_schema.yml", package = "austraits.build")`. 

AusTraits is essentially a series of linked components, which cross link against each other::

```{r, results="hide", comment = '', eval=FALSE}
names(austraits) %>%
  create_tree_branch("austraits.build") %>%
  writeLines()
```

```
austraits
├── traits
├── locations
├── contexts
├── methods
├── excluded_data
├── taxonomic_updates
├── taxa
├── contributors
├── sources
├── definitions
├── schema
├── metadata
└── build_info
```

These include all the data and contextual information submitted with each contributed dataset. 

# Components

The core components are defined as follows.

```{r}
print_schema_element <- function(elements) {
  if (elements$type == "character") {
    sprintf("**Content:** %s\n", elements$value) %>%
      writeLines()
  }

  if (elements$type == "table") {
    sprintf("**Content:** \n") %>%
      writeLines()

    elements$elements %>%
      util_list_to_df1() %>%
      my_kable_styling() %>%
      writeLines()
  }
}
```

## traits {#traits}
```{r}
elements <- schema$austraits$elements$traits

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

**Details:**  

#### Entity type

An entity is the `feature of interest`, indicating what a trait value applies to. While an entity can be just a component of an organism, within the scope of AusTraits, an `individual` is the finest scale entity that can be documented. The same study might measure some traits at a population-level (`entity = population`) and others at an individual-level (`entity = individual`).

In detail:

- `entity_type` is `r tolower(schema$metadata$elements$traits$elements$entity_type)` Possible values are: 

```{r entity_type}
schema$entity_type$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

#### Identifiers

The traits table includes 12 identifiers, `dataset_id`, `observation_id`, `taxon_name`, `population_id`, `individual_id`, `temporal_id`, `source_id`, `location_id`, `entity_context_id`, `plot_id`, `treatment_id`, and `method_id`.

`dataset_id`, `source_id` and `taxon_name` have easy-to-interpret values. The others are simply integral identifiers that link groups of measurements and are automatically generated through the AusTraits workflow (`individual_id` can be assigned in the metadata file or automatically generated.)

To expand on the definitions provided above,

- `observation_id` links measurements made on the same entity (individual, population, or species) at a single point in time.

- `population_id` indicates entites that share a common `location_id`, `plot_id`, and `treatment_id`. It is used to align measurements and observation_id's for `individuals` versus `populations` (i.e. distinct `entity_types`) that share a common `population_id`. It is numbered sequentially within a dataset.

- `individual_id` indicates a unique organisms. It is numbered sequentially within a dataset by population. Multiple observations on the same organism across time (with distinct `observation_id` values), share a common `individual_id`.

- `temporal_id` indicates a distinct point in time and is used only if there are repeat measurements on a population or individual across time. The identifier links to context properties (& their associated information) in the `contexts` table for context properties of type `temporal`.

- `source_id` is applied if not all data within a single dataset (`dataset_id`) is from the same source, such as when a dataset represents a compilation for a meta-analysis.

- `location_id` links to a distinct `location_name` and associated `location_properties` in the `location` table.

- `entity_context_id` links to information in the `contexts` table for context properties (& associated values/descriptions) with category `entity_context`. 
`Entity_contexts` include organism sex, organism caste and any other features of an entity that needs to be documented.

- `plot_id` links to information in the `contexts` table for context properties (& associated values/descriptions) with category `plot`. `Plot contexts` include both blocks/plots within an experimental design as well as any stratified variation within a location that needs to be documented (e.g. slope position).

- `treatment_id`links to information in the `contexts` table for context properties (& associated values/descriptions) with category `treatment`. `Treatment contexts` are experimental manipulations applied to groups of individuals.

- `method_id`links to information in the `contexts` table for context properties (& associated values/descriptions) with category `method`. A `method context` indicates that the same trait was measured on or across individuals using different methods.

As well, `measurement_remarks` is used to document `r tolower(schema$metadata$elements$traits$elements$measurement_remarks)`

#### Life stage, basis of record 

- `life_stage` is `r tolower(schema$metadata$elements$traits$elements$life_stage)`.

- `basis_of_record` is `r tolower(schema$metadata$elements$traits$elements$basis_of_record)` Possible values are:

```{r basis_of_record}
schema$basis_of_record$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

#### Values, Value types, basis of value {#value_types}

Each record in the table of trait data has an associated `value`,  `value_type`, and `basis_of_value`. 

`Value type:` A trait's `value_type` is either `numeric` or `categorical`. 
  - For traits with numerical values, the recorded value has been converted into standardised units and the AusTraits workflow has confirmed the value can be converted into a number and lies within the allowable range. 
  - For categorical variables, records have been aligned through substitutions to values listed as allowable values (terms) in a trait's definition. 
      * we use `_` for multi-word terms, e.g. `semi_deciduous`  
      * we use a space for situations where two values co-occur for the same entity. For instance, a flora might indicate that a plant species can be either annual or biennial, in which case the trait is scored as `annual biennial`.

Each trait measurement has an associated `value_type`, which is `r tolower(schema$metadata$elements$traits$elements$value_type)` Possible values are:

```{r value_type}
schema$value_type$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

Each trait measurement also has an associated `basis_of_value`, which is `r tolower(schema$metadata$elements$traits$elements$basis_of_value)` Possible values are: 

```{r basis_of_value}
schema$basis_of_value$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```


AusTraits does not include intra-individual observations made at a single point in time. When multiple measurements per individual are submitted to AusTraits, we take the mean of the values and record the value_type as `mean` and indicate under replicates the number of measurements made.  




## locations
```{r}
elements <- schema$austraits$elements$locations

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

## contexts
```{r}
elements <- schema$austraits$elements$contexts

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

## methods {#methods}
```{r}
elements <- schema$austraits$elements$methods

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## exluded_data
```{r}
elements <- schema$austraits$elements$excluded_data

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## taxa

**Description:** A table containing details on taxa that are included in the table [`traits`](#traits). We have attempted to align species names with known taxonomic units in the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc) and/or the [`Australian Plant Names Index` (APNI)](https://biodiversity.org.au/nsl/services/APNI); the sourced information is released under a CC-BY3 license. 


Version `r desc::desc_get_field("Version")` of AusTraits contains records for `r austraits$traits$taxon_name %>% unique() %>% length()` different taxa. 


```{r}
elements <- schema$austraits$elements$taxa

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## taxonomic_updates

```{r}
elements <- schema$austraits$elements$taxonomic_updates

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


Both the original and the updated taxon names are included in the [`traits`](#traits) table.


## definitions
```{r}
elements <- schema$austraits$elements$definitions

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```

**Details on trait definitions:** The allowable trait names and trait values are defined in the definitions file. Each trait is 
labelled as either `numeric` or `categorical`. An example of each type is as follows. For the full list, see the [Trait definitions vignette](http://traitecoevo.github.io/austraits.build/articles/trait_definitions.html).


```{r, traits}
for (trait in c("leaf_mass_per_area", "woodiness")) {
  elements <- austraits$schema$traits$elements[[trait]]

  data_trait <- austraits$traits %>% filter(trait_name == trait)

  c(sprintf("**%s**\n\n", trait), sprintf("- label: %s", elements$label), sprintf("- description: %s", elements$description), sprintf("- number of records: %s", data_trait %>% nrow()), sprintf("- number of studies: %s", data_trait %>% pull(dataset_id) %>% unique() %>% length()), sprintf("- type: %s%s", elements$type, ifelse(elements$type == "numeric", sprintf("\n- units: %s", elements$units), "")), ifelse(elements$type == "numeric", sprintf("- allowable range: %s - %s %s", elements$values$minimum, elements$values$maximum, elements$units), sprintf("- allowable values:\n%s\n", paste0("    - *", elements$values %>% names(), "*: ", elements$values %>% unlist(), collapse = "\n"))), "") %>% writeLines()
}
```


## contributors
```{r}
elements <- schema$austraits$elements$contributors

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```


## sources

For each dataset in the compilation there is the option to list primary and secondary citations. The primary citation is defined as, ``r austraits$schema$metadata$elements$source$values$primary$description`` The secondary citation is defined as, ``r austraits$schema$metadata$elements$source$values$secondary$description`` 

The element `sources` includes bibtex versions of all sources which can be imported into your reference library:

```
RefManageR::WriteBib(austraits$sources, "refs.bib") # write all sources to file
RefManageR::WriteBib(austraits$sources["Falster_2005_1"], "refs.bib") # write a single reference to a file
```

Or individually viewed:
```
austraits$sources["Falster_2005_1"]
```

A formatted version of the sources also exists within the table [methods](#methods). 


## metadata
```{r}
elements <- schema$austraits$elements$metadata

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  writeLines(c(""))
```


## build_info
```{r}
elements <- schema$austraits$elements$build_info

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_schema_element()

  writeLines(c(""))
```
