---
title: Structure of AusTraits data compilation
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ Structure of AusTraits data compilation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = "asis",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(austraits.build)
library(dplyr)
library(readr)

root.dir <- rprojroot::find_root("remake.yml")
knitr::opts_knit$set(root.dir = root.dir)

my_kable_styling <- my_kable_styling_html
```

```{r, echo=FALSE, results='hide', message=FALSE}
## Loads austraits into global name space
  austraits <- remake::make("austraits")
  definitions <- austraits$definitions
```


This document describes the structure of the AusTraits compilation, corresponding to **Version `r desc::desc_get_version("./DESCRIPTION")`** of the dataset. 

Note that the information provided below is based on the information provided within the file `definitions.yml`. 

AusTraits is essentially a series of linked components, which cross link against each other::

```{r, results="hide", comment = '', eval=FALSE}
names(austraits) %>%
  create_tree_branch("austraits") %>%
  writeLines()
```

```
austraits
├── traits
├── sites
├── contexts
├── methods
├── excluded_data
├── taxa
├── taxonomic_updates
├── definitions
├── contributors
├── sources
└── build_info
```

These include all the data and contextual information submitted with each contributed dataset. 

# Components

The core components are defined as follows.

```{r}
print_definitions_element <- function(elements) {
  if (elements$type == "character") {
    sprintf("**Content:** %s\n", elements$value) %>%
      writeLines()
  }

  if (elements$type == "table") {
    sprintf("**Content:** \n") %>%
      writeLines()

    elements$elements %>%
      list1_to_df() %>%
      my_kable_styling() %>%
      writeLines()
  }
}
```

## traits {#traits}
```{r}
elements <- austraits$definitions$austraits$elements$traits

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```

**Details:**  

#### Observation IDs

Each trait measurement has an associated `observation_id`.  Observation IDs bind together related measurements within any dataset, and thereby allow transformation between long format (e.g. with variables `trait_name` and `value`) and wide format (e.g. with traits as columns). When multiple traits were collected on the same individual, the `observation_id` identified which measurements were made on the same individual (or population,  species depending on the level of detail supplied). Generally, `observation_id` has the format `dataset_id_XX` where `XX` is a unique number within each dataset. 

For datasets submitted to AusTraits in long format, a column specifying `observation_id` should already exist, assigned based on a specified grouping variable.  If missing, `observation_id` is assigned based on `species_name`. 

For datasets submitted to AusTraits in wide format each row is assigned a unique `observation_id` during processing, allowing conversion to long format. 


#### Values and Value types {#value_types}

Each record in the table of trait data has an associated `value` and `value_type`. 

Traits are either `numeric` or `categorical`. For traits with numerical values, the recorded value has been converted into standardised units and we have checked that the value can be converted into a number and lies within the allowable range.  For categorical variables, we only include records whose values are included as allowable values in a trait's definition in the definitions table. Moreover, we use a format whereby

- we use `_` for multi-word terms, e.g. `semi_deciduous`  
- we use a space for situations where a single study reports two possible values for that trait, e.g. `annual biennial` for something which is either annual or biennial 

Each trait measurement also has an associated `value_type`, which is ``r austraits$definitions$value_type$description`` Possible values are:  

```{r value_type}
austraits$definitions$value_type$values %>%
  list1_to_df() %>%
  my_kable_styling() %>%
  writeLines()
```

AusTraits does not include intra-individual observations. When multiple measurements per individual are submitted to AusTraits, we take the mean of the values and record the value_type as `individual_mean`.  

## sites
```{r}
elements <- austraits$definitions$austraits$elements$sites

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```

## contexts
```{r}
elements <- austraits$definitions$austraits$elements$contexts

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```

## methods {#methods}
```{r}
elements <- austraits$definitions$austraits$elements$methods

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```


## exluded_data
```{r}
elements <- austraits$definitions$austraits$elements$excluded_data

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```


## taxa

**Description:** A table containing details on taxa that are included in the table [`traits`](#traits). We have attempted to align species names with known taxonomic units in the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc) and/or the [`Australian Plant Names Index` (APNI)](https://biodiversity.org.au/nsl/services/APNI); the sourced information is released under a CC-BY3 license. 


Version `r desc::desc_get_field("Version")` of AusTraits contains records for `r austraits$traits$taxon_name %>% unique() %>% length()` different taxa. 


```{r}
elements <- austraits$definitions$austraits$elements$taxa

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```


## taxonomic_updates

```{r}
elements <- austraits$definitions$austraits$elements$taxonomic_updates

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```


Both the original and the updated taxon names are included in the [`traits`](#traits) table.


## definitions
```{r}
elements <- austraits$definitions$austraits$elements$definitions

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```

**Details on trait definitions:** The allowable trait names and trait values are defined in the definitions file. Each trait is 
labelled as either `numeric` or `categorical`. An example of each type is as follows. For the full list, see the section on Trait definitions.


```{r, traits}
for (trait in c("specific_leaf_area", "woodiness")) {
  elements <- austraits$definitions$traits$elements[[trait]]

  data_trait <- austraits$traits %>% filter(trait_name == trait)

  c(sprintf("**%s**\n\n", trait), sprintf("- label: %s", elements$label), sprintf("- description: %s", elements$description), sprintf("- number of records: %s", data_trait %>% nrow()), sprintf("- number of studies: %s", data_trait %>% pull(dataset_id) %>% unique() %>% length()), sprintf("- type: %s%s", elements$type, ifelse(elements$type == "numeric", sprintf("\n- units: %s", elements$units), "")), ifelse(elements$type == "numeric", sprintf("- allowable range: %s - %s %s", elements$values$minimum, elements$values$maximum, elements$units), sprintf("- allowable values:\n%s\n", paste0("    - *", elements$values %>% names(), "*: ", elements$values %>% unlist(), collapse = "\n"))), "") %>% writeLines()
}
```


## contributors
```{r}
elements <- austraits$definitions$austraits$elements$contributors

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```


## sources

For each dataset in the compilation there is the option to list primary and secondary citations. The primary citation is defined as, ``r austraits$definitions$metadata$elements$source$values$primary`` The secondary citation is defined as, ``r austraits$definitions$metadata$elements$source$values$secondary`` 

The element `sources` includes bibtex versions of all sources which can be imported into your reference library:
```
WriteBib(austraits$sources)                             #write all sources to file
WriteBib(austraits$sources["Falster_2005_1"])           #write a single reference to a file
```

Or individually viewed:
```
austraits$sources["Falster_2005_1"]
```

A formatted version of the sources also exists within the table [methods](#methods). 



## build_info
```{r}
elements <- austraits$definitions$austraits$elements$build_info

  sprintf("**Description:** %s.\n", elements$description) %>%
    stringr::str_replace_all("\\.\\.", "\\.") %>%
    writeLines()

  elements %>%
    print_definitions_element()

  writeLines(c(""))
```
