---
title: Report on study from the `Austraits` data compilation
output:
  html_document:
    df_print: kable
    highlight: tango
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# knitr defaults
root.dir = rprojroot::find_root("remake.yml")
knitr::opts_knit$set(root.dir = root.dir)
knitr::opts_chunk$set(echo=FALSE, cache=FALSE)
# default for table format
options(knitr.table.format = "html")

# Guidelines for writing report code
# - use tidyverse style and format: http://htmlpreview.github.io/?https://github.com/nicercode/2018_BEES_regression/blob/master/tidyverse.html
# - use kableExtra for styling: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# - use knitr chunck options: https://rmarkdown.rstudio.com/lesson-3.html

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
library(knitr)
library(dplyr)
library(leaflet)
library(ggbeeswarm)
library(ggbeeswarm)
library(kableExtra)

source("R/austraits.R")
source("R/support.R")
source("R/report_utils.R")
source("R/report_notetaker.R")

## Assumes to items exist in global name space
if(!exists("austraits")) {
  austraits <- remake::make("austraits")
  # stop("austraits must exist in global name space to knit a report")
}
if(!exists("dataset_id")) {
  dataset_id <- "Blackman_2014"
  #  stop("dataset_id must exist in global name space to knit a report")
}

definitions <- austraits$definitions
data_study <- extract_dataset(austraits, dataset_id)
data_yaml <- read_yaml(sprintf("data/%s/metadata.yml", dataset_id))
metadata_study <- austraits$metadata[[dataset_id]]

knitr::opts_chunk$set(fig.cap='', fig.path = file.path(root.dir,"export/figures", paste0(dataset_id,"_")))

# start notetaker device
questions <- start_notetaker()

new_question <- function(txt) {
  x <- questions(add_note, as_note(txt)) %>% print_notes(as_anchor=TRUE)
  writeLines(sprintf("<font color='orange'>**Question:** %s</font>\n\n", paste0(x, collape="")))
}
```

# Introduction

## About Austraits

`Austraits` is an open-source compilation of trait data for Australian plant species. The traits we are considering are any structural or physiological variable that: 1. Tends to vary widely among species; and 2. Have been collected across a moderate number of species. Many people collect trait data, and as such, data is spread across many sources, under a variety of formats and terms. 

The goal of the Austraits project is to combine all of these data fragments into a harmonised compilation, to be made publicly available.

Austraits is built from many different sources, here referred to as `studies` (so called because the sources are most-often individual scientific papers), with each study denoted by a distinct `dataset_id`. Our data processing pipeline seeks to combine all of the different studies in a transparent and reproducible way. In addition, the Austraits repository consists of configuration files defining the list of known species, the definitions for each trait and table, and appropriate unit conversions.

## Purpose of this report

The Austraits compilation includes data from a study with dataset_id **` `r dataset_id` `**. This document describes the information we have on that study.
This report gives both you, as a data contributor, and us, as Austraits compilers, the opportunity to:

1. Verify that our understanding and handling of your data is correct
2. Provide any missing information about the data

Data from each study in Austraits is organised into two files, within its own folder:

- ` data/`r dataset_id`/data.csv`: contains the primary trait measurements, in either long or wide format
- ` data/`r dataset_id`/metadata.yml`: contains all the contextual data (or metadata) about the trait measurements.

Data from those files has been used to generate the report below.

Please review the report and the original files, check they are correct, and send updates where appropriate. Please also answer the queries in the report below, appearing in orange. These appear again as a list at the end of the document.

We would welcome any of the following:

- Answers to the list of questions
- Expanding or correcting the information in the `metadata.yml`
- Updates to the data

To aid your review, we have attached the following files:

- `Austraits.html`: A document describing the Austraits compilation
- `Workflow`: A document describing the data structures and workflow used to build `Austraits`
- ` `r dataset_id`/data.csv` and ` `r dataset_id`/metadata.yml`: The raw inputs for your study.

Send feedback by replying to the email XXXXX, attaching:

1. Answers to the questions we ask throughout the document, and summarised at the end of the document.
2. Edited versions of the input files for your study: ` `r dataset_id`/data.csv` and ` `r dataset_id`/metadata.yml` (if appropriate).

# Study context

We collect information about the context of the trait measurements. This includes details on the people involved, the sources, and details about type of collection, including location and time of collections. 

## People

The following people are listed as contacts for this study:

```{r, results='asis', echo=FALSE}
metadata_study$people %>%
  list_to_df() %>%
  select(role, name, institution) %>%
  my_kable_styling()
```

Institutions are optional. Roles are as follows:

- **collector**: The person (people) leading data collection (up to 2 permitted)
- **contributor**: Person responsible for entering data into Austraits
- **lab_leader**: Leader of lab group at time of collection
- **assistant**: Anyone else who assisted in collection of the data

```{r, results='asis', echo=FALSE}
new_question("(section `people`) Are all appropriate people listed, with appropriate details?")
```

## Sources

Data from this study is recorded as coming from the following sources:

Primary citation:

```{r, results='asis', echo=FALSE}
metadata_study$source$primary %>%
  list1_to_df() %>%
  my_kable_styling()
```

```{r, results='asis', echo=FALSE}
if(!is.null(metadata_study$source$secondary)) {
  writeLines("Secondary citation:")
  metadata_study$source$secondary %>%
    list1_to_df() %>%
    my_kable_styling()
  }
```


```{r, results='asis', echo=FALSE}
new_question("(section `source`) Are the citation details for this study correct?")
```

## General description

The description we have of the study is as follows:

```{r, results='asis', echo=FALSE}
metadata_study$dataset %>%
  list1_to_df() %>%
  my_kable_styling()
```

```{r, results='asis', echo=FALSE}
new_question("(section `dataset`) Can you provide more detailed information on the study origin and timing?")
```

## Locations sampled

Data were collected at the following sites and locations. As a minimum, we are aiming to collect, the variables

```{r, results='asis', echo=FALSE, warning=FALSE}
site_main <- c("longitude (deg)","latitude (deg)", "description")
sites <- data_study$context %>%
  filter(trait_name %in% site_main) %>%
  select(-dataset_id) %>%
  spread(trait_name, value) %>% 
  mutate(description = if (exists('description', where=.)) description else NA)
  

if(nrow(sites) > 0 ) {
  sites <- sites %>%
  mutate(`latitude (deg)` = as.numeric(`latitude (deg)`),
         `longitude (deg)` = as.numeric(`longitude (deg)`)) %>%
  select(site_name, `latitude (deg)`, `longitude (deg)`, description)

  sites %>%
    my_kable_styling()
} else{
    new_question("(section `sites`) **No location data available!!** Can you provide location details where your data were sampled?")
}

```

```{r maps, eval= TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.keep="all"}
# Make a map of study sites using leaflet
sites <- sites %>% filter(!is.na(`latitude (deg)`) & !is.na(`longitude (deg)`)) 
if(nrow(sites) > 0 ) {
  leaflet() %>% 
  addTiles() %>%
  addScaleBar() %>% 
  addMiniMap(zoomLevelFixed=2) %>% 
  addMarkers(lng = sites$`longitude (deg)`, lat = sites$`latitude (deg)`,
             label = sites$site_name)	
}
```

```{r, results='asis', echo=FALSE}
new_question("(section `sites`) Do site details look complete? As a minimum we would like `latitude`, `longitude`, `description`.")
```

## Additional site data (optional)

The following variables have been recorded for the sites included in the study:

```{r, results='asis', echo=FALSE}

data_study$context %>%
  filter(!(trait_name %in% site_main)) %>%
  select(-dataset_id) %>%
  spread(trait_name, value) -> context

if(nrow(context) > 0 ) {
  context %>%
    my_kable_styling()

  new_question("(section `sites`) Do the additional contextual site details look complete?")
  } else{
  writeLines("No contextual site data available.")
}

```

# Trait measurements

The dataset includes `r data_study$data %>% nrow()` individual data points from `r data_study$data$species_name %>% unique() %>% length()` species, with data included for the following `r data_study$data$trait_name %>% unique() %>% length()` traits:

```{r summary_table, results='asis', echo=FALSE}
n_records <- data_study$data %>%
  group_by(trait_name) %>%
    summarise(
        species = length(unique(species_name)),
        records = length(value)) %>%
   mutate(records_per_species = round(records / species, digits=2) ) 

n_records %>%
  my_kable_styling()
```


```{r, results='asis', echo=FALSE}
new_question("(section `traits`) Does this study include other trait data we may have missed?")

if(median(n_records$records_per_species) < 2) {
  new_question("(section `traits`) It appears as though we mostly have means for each species. Can you provide individual-level measurements?")
}

new_question("(section `traits`) Were any of your data sourced from other studies? If so, can you tell us which records and the source (so that we can avoid duplicates, where possible)?")

```

```{r excluded,  results='asis', echo=FALSE}

writeLines(ifelse(nrow(data_study$excluded_data) == 0,
        "All data passed quality control so nothing was excluded in this study. (Yay!)",
        "We excluded some records in your dataset as they did not pass the quality control. The number of points and reasons for exclusion are as follows:"))

if(nrow(data_study$excluded_data) > 0 ) {
  data_study$excluded_data %>% 
    group_by(trait_name, error) %>%
    summarise(`points excluded` = n())  %>%
    my_kable_styling() %>% 
    writeLines()

  new_question("(section `traits`) Can you provide any additional information so that above exclusions no longer apply?")
}
```


## Numerical traits 

```{r numerical, results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10}
traits <- data_study$data$trait_name %>% unique() %>% sort()
traits_numeric <- traits[trait_is_numeric(traits, definitions)]

writeLines(ifelse(length(traits_numeric) == 0,
        "No numerical traits data are currently available in this dataset",
        "We have recorded the following traits with continuous numerical values in this dataset. Plots are shown comparing the distribution of values in your study (blue) to those in other datasets within austraits (red, green). For each trait, please eyeball the information provided and let us know of any issues or updates where appropriate."))

for(trait in traits_numeric) {


  data_trait_study <- data_study$data %>%
    filter(trait_name == trait) %>%
    mutate(value = as.numeric(value))

  data_trait_all <- austraits$data %>%
    filter(trait_name == trait) %>%
    mutate(value = as.numeric(value))

  transforms <- data_yaml$traits %>% list_to_df() %>% filter(trait_name == trait)

  # Adjust height of plot according number of rows?
  # n_study <- data_trait_all %>% dplyr::pull(dataset_id) %>% unique() %>% length()
  # figure_height <- 4 + max(12, (n_study/5))

  c("", "",
    sprintf("## %s", definitions$traits$values[[trait]]$label),
    "",
    sprintf("We aligned the variable called `%s`in the data you supplied with the trait `%s` in the Austraits database.\n\nThe original variable was supplied with units `%s`; these were converted to our standard `%s`.\n\nThe contributed data has the following properties:", transforms$var_in, trait, transforms$unit_in, definitions$traits$values[[trait]]$units),
    "",
    sprintf("- **Standardised name**: %s", trait ),
    sprintf("- **Standardised description**: %s", definitions$traits$values[[trait]]$description ),
    sprintf("- **Records**: %s in this study (of %s in Austraits)", data_trait_study %>% nrow(), data_trait_all %>% nrow() ),
    sprintf("- **Allowable range**: %s - %s", definitions$traits$values[[trait]]$values$minimum,
            definitions$traits$values[[trait]]$values$maximum ),
    sprintf("- **Observed range in this study**: %s - %s", data_trait_study %>% dplyr::pull(value) %>% min(),
            data_trait_study %>% dplyr::pull(value) %>% max() ),
    "", "",
    "Data for this trait in this study were collected using the following methods:", "",
    sprintf("- **Value type**: %s", transforms$value_type),
    sprintf("- **Replicates**: %s", transforms$replicates),
    sprintf("- **Methods (from paper)**: %s", transforms$methods), 
    "","") %>%
  writeLines()
  
  trait_distribution_plot_numerical(austraits, trait, "dataset_id", highlight=dataset_id, hide_ids = TRUE)

  writeLines(c(""))
  
  methods_missing <- transforms[1,] %in% c("unknown", NA)

  new_question(
    sprintf("(section `traits`) Do the methods data data for the trait appear %s correct? %s", trait, 
            ifelse(any(methods_missing), 
                   sprintf("Can you provide missing details for trait `%s`: %s",trait, paste0(names(transforms)[methods_missing], collapse= ", ")), 
                   "")
            )
    )
  }
```

## Categorical traits

```{r categorical, results='asis', echo=FALSE, eval=TRUE}

traits <- data_study$data$trait_name %>% unique() %>% sort()
traits_categorical <- traits[!trait_is_numeric(traits, definitions)]

writeLines(
  ifelse(length(traits_categorical) == 0,
    "No categorical traits data are currently available in this dataset.",
    "We have recorded the following traits with categorical values in this dataset. Plots are shown comparing the distribution of values in your study to others in the dataset"))

for(trait in traits_categorical) {

  transforms <- data_yaml$traits %>% list_to_df() %>% filter(trait_name == trait)

  data_trait_study <- data_study$data %>%
    filter(trait_name == trait)

  data_trait_all <- austraits$data %>%
    filter(trait_name == trait)

  x <- c("", "",
    sprintf("### %s", definitions$traits$values[[trait]]$label),
    "",
    sprintf("We aligned the variable called `%s`in the data you supplied with the trait `%s` in the Austraits database.\n\nThe contributed data has the following properties:", transforms$var_in, trait),
    "",
    sprintf("- **Standardised name**: %s", trait ),
    sprintf("- **Standardised description**: %s", definitions$traits$values[[trait]]$description ),
#    sprintf("- **Records**: %s in this study (of %s in Austraits)", data_trait_study %>% nrow(), data_trait_all %>% nrow() ),
#    sprintf("- **Observed values**: %s of %s allowed values recorded -- %s", length(actual),length(allowed), actual %>% paste0(collapse=", ") %>% sort() ),
    "",
    "The following table shows distribution of recorded values in your study and Austraits as a whole:",
    "")

  writeLines(x)

  right_join(
    data_trait_study %>% group_by(value) %>% summarise(study = n()),
    data_trait_all %>% group_by(value) %>% summarise(total = n()),
    by= "value"
    ) %>% 
  mutate(
    study = ifelse(is.na(study), 0, study),
    `% study` = round(study/sum(study)*100),
    `% total` = round(total/sum(total)*100)
  ) %>%
  my_kable_styling() %>% 
  add_header_above(c(" ", "Counts" = 2, "Percent" = 2)) %>%
  column_spec(1, border_right = TRUE) %>%
  column_spec(3, border_right = TRUE) %>%
  column_spec(5, border_right = TRUE) %>%
  writeLines()
}
```


# Species taxonomy

We have records on the following species from your study. We have attempted to align species names with known taxonomic units, focussing primarily on the [`The Plant List` (TPL)](http://www.theplantlist.org/) -- a global working list of all known plant species. In addition we have tried to align these names with the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc) and the [`Australian Plant Names Index` (APC_name_ID)](https://biodiversity.org.au/nsl/services/APNI).

The table below shows the aligned taxonomy, including

- the original name supplied to us 
- alignment of your species with the [`The Plant List` (TPL)](http://www.theplantlist.org/)
- alignment of your species with the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc)
- alignment of your species with the [`Australian Plant Names Index` (APC_name_ID)](https://biodiversity.org.au/nsl/services/APNI)
- the reason a name change was implemented (where relevant) 

Updated names are shown only where different to the original name supplied. We also include links to the species records in the above  systems. Clicking on the links will take you to the relevant taxonomic record in each system for that species.

```{r species_list, results='asis', echo=FALSE}

corrections <- tibble(original_name= character(), notes = character())

if(!is.null(data_yaml$taxonomic_updates)) {
  corrections <- data_yaml$taxonomic_updates %>%
    list_to_df() %>% 
    select(original_name = find, notes = reason) 
}

species <- data_study$data %>%
  select(original_name, species_name) %>%
  distinct() %>%
  arrange(original_name) %>%
  left_join(austraits$species_list, by ="species_name") %>% 
  rename(TPL_name = species_name) %>% 
  mutate(TPL_name = ifelse(TPL_name == original_name, "", TPL_name),
         APC_name = ifelse(APC_name == original_name, "", APC_name)
         ) %>%
 left_join(corrections, by = "original_name")  %>%
 mutate(notes = ifelse(is.na(notes), "", notes))
 

species %>%
  mutate(
    TPL_ID = as_link_tpl(TPL_ID),
    APC_ID = as_link_apc(APC_ID),
    APNI_ID = as_link_apni(APNI_ID)
  ) %>%
  my_kable_styling() 
 # borders -- seems to be too slow on large datasets
 # %>%
 #  kableExtra::column_spec(1, border_right = TRUE) %>%
 # kableExtra::column_spec(6, border_right = TRUE) %>%
 #  kableExtra::column_spec(8, border_right = TRUE) %>%
 # kableExtra::column_spec(9, border_right = TRUE)

```

```{r, results='asis', echo=FALSE}
new_question("(section `taxonomic_updates`) Do these taxonomic alignments and corrections look reasonable?")
```

# Questions

Here is a summary of the questions we would like answers to. Clicking on the links will take you back to the appropriate spot in the document. Please copy and past these into an email when you respond. 

<font size = '3' color='orange'>


```{r, results='asis', echo=FALSE}
questions() %>% print_all_notes() %>% writeLines()
```


</font>


# Session information

The technical information below may useful for us for the sake of reproducibility, but you don't need to worry about it!

<font color='grey'>
This report was generated using the data from the commit `r get_SHA_link()`

And the following R environment

```{r, echo=FALSE}
sessionInfo()
```

</font>

