---
title: "Dictionary for `AusTraits` data compilation - a curated plant trait database for the Australian flora"
output:
  html_document:
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: true
      smooth_scroll: true
params:
  austraits: provide
editor_options:
  chunk_output_type: console
---

<!-- hack to get indentation on 3rd level of floating TOC; see
https://stackoverflow.com/questions/46201753/rmarkdown-indentation-of-toc-items-in-html-output
 -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# knitr defaults
knitr::opts_chunk$set(echo=FALSE, cache=FALSE, results='asis', message=FALSE, warning=FALSE)

# default for table format
options(knitr.table.format = "html")

# Guidelines for writing report code
# - use tidyverse style and format: http://htmlpreview.github.io/?https://github.com/nicercode/2018_BEES_regression/blob/master/tidyverse.html
# - use kableExtra for styling: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# - use knitr chunk options: https://rmarkdown.rstudio.com/lesson-3.html

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
library(tidyverse)
library(RefManageR)
library(knitr)
library(kableExtra)

my_kable_styling <- util_kable_styling_html

definitions <- austraits$definitions
schema <- austraits$schema
```

This document describes the structure of the AusTraits compilation, corresponding to **Version `r austraits$build_info$version`** of the dataset. 
Note that the information provided below is based on the information provided within the file `system.file("data", "schema.yml", package = "austraits.build")`. 

For details on access, structure and usage please visit https://doi.org/10.5281/zenodo.3568417

# Elements of AusTraits

The compiled AusTraits database has the following main components:

```
austraits
├── traits
├── locations
├── contexts
├── methods
├── excluded_data
├── taxonomic_updates
├── taxa
├── contributors
├── sources
├── definitions
├── schema
├── metadata
└── build_info
```

These elements include all the data and contextual information submitted with each contributed datasets. Each component  is defined as follows:

```{r}
print_definitions_element <- function(elements) {
  if(elements$type == "character") {
    sprintf("**Content:** %s\n", elements$value) %>% 
      writeLines()
  }
  
  if(elements$type == "table") {
    
    sprintf("**Content:** \n") %>% 
      writeLines()

    elements$elements %>% 
      util_list_to_df1() %>%
      kableExtra::kable() %>% 
      writeLines()
  }
}

for(v in names(austraits)) {
  elements <- austraits$schema$austraits$elements[[v]]
  
  sprintf("## %s {#%s}\n\n**Description:** %s\n", v, v, elements$description) %>% 
    writeLines()
  
  elements %>%
    print_definitions_element()

  writeLines(c(""))
}
```


# Methological details

## Dataset IDs

The core organising unit behind AusTraits is the `dataset_id`. Records are organisation as coming from a particular study, defined by the  `dataset_id`. Our preferred format for `dataset_id` is surname of the first author of any corresponding publication, followed by the year, as `surname_year`. E.g. `Falster_2005`. Wherever there are multiple studies with the same id, we add a suffix `_2`, `_3` etc. E.g.`Falster_2005`, `Falster_2005_2`.

## Other IDs

As well as a `dataset_id`, each trait measurement has 10 additional identifiers, `observation_id`, `population_id`, `individual_id`, `temporal_id`, `source_id`, `location_id`, `entity_context_id`, `plot_id`, `treatment_id`, and `method_id`.

All except source_id are  simply integral identifiers that link groups of measurements and are automatically generated through the AusTraits workflow (`individual_id` can be assigned in the metadata file or automatically generated.)

To expand on the definitions provided above,

- `observation_id` links measurements made on the same entity (individual, population, or species) at a single point in time.

- `population_id` indicates entites that share a common `location_id`, `plot_id`, and `treatment_id`. It is used to align measurements and observation_id's for `individuals` versus `populations` (i.e. distinct `entity_types`) that share a common `population_id`. It is numbered sequentially within a dataset.

- `individual_id` indicates a unique organisms. It is numbered sequentially within a dataset by population. Multiple observations on the same organism across time (with distinct `observation_id` values), share a common `individual_id`.

- `temporal_id` indicates a distinct point in time and is used only if there are repeat measurements on a population or individual across time. The identifier links to context properties (& their associated information) in the `contexts` table for context properties of type `temporal`.

- `source_id` is applied if not all data within a single dataset (`dataset_id`) is from the same source, such as when a dataset represents a compilation for a meta-analysis.

- `location_id` links to a distinct `location_name` and associated `location_properties` in the `location` table.

- `entity_context_id` links to information in the `contexts` table for context properties (& associated values/descriptions) with category `entity_context`. 
`Entity_contexts` include organism sex, organism caste and any other features of an entity that needs to be documented.

- `plot_id` links to information in the `contexts` table for context properties (& associated values/descriptions) with category `plot`. `Plot contexts` include both blocks/plots within an experimental design as well as any stratified variation within a location that needs to be documented (e.g. slope position).

- `treatment_id`links to information in the `contexts` table for context properties (& associated values/descriptions) with category `treatment`. `Treatment contexts` are experimental manipulations applied to groups of individuals.

- `method_id`links to information in the `contexts` table for context properties (& associated values/descriptions) with category `method`. A `method context` indicates that the same trait was measured on or across individuals using different methods.

## Values, Value types, basis of value {#value_types}

Each record in the table of trait data has an associated `value`,  `value_type`, and `basis_of_value`. 

`Value type:` A trait's `value_type` is either `numeric` or `categorical`. 
  - For traits with numerical values, the recorded value has been converted into standardised units and the AusTraits workflow has confirmed the value can be converted into a number and lies within the allowable range. 
  - For categorical variables, records have been aligned through substitutions to values listed as allowable values (terms) in a trait's definition. 
      * we use `_` for multi-word terms, e.g. `semi_deciduous`  
      * we use a space for situations where two values co-occur for the same entity. For instance, a flora might indicate that a plant species can be either annual or biennial, in which case the trait is scored as `annual biennial`.

Each trait measurement has an associated `value_type`, which is `r tolower(schema$metadata$elements$traits$elements$value_type)` Possible values are:

```{r value_type}
schema$value_type$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

Each trait measurement also has an associated `basis_of_value`, which is `r tolower(schema$metadata$elements$traits$elements$basis_of_value)` Possible values are: 

```{r basis_of_value}
schema$basis_of_value$values %>%
  util_list_to_df1() %>%
  my_kable_styling() %>%
  writeLines()
```

AusTraits does not include intra-individual observations made at a single point in time. When multiple measurements per individual are submitted to AusTraits, we take the mean of the values and record the value_type as `mean` and indicate under replicates the number of measurements made.  

## Taxonomy

Version `r austraits$build_info$version` of AusTraits contains records for `r austraits$taxa %>% nrow()` different taxa. We have aligned taxa with known taxonomic units in the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc) and/or the [`Australian Plant Names Index` (APNI)](https://biodiversity.org.au/nsl/services/APNI). Of the `r austraits$taxa %>% nrow()` taxa included, `r austraits$taxa %>% filter(!is.na(taxonomic_reference)) %>% nrow()` are aligned with known taxa. 

The `traits` table reports both the original and the updated taxon name alongside each trait record.

The table `taxa` lists all taxa in the database, including additional information about the taxa. 

The table `taxonomic_updates` provides details on all taxonomic names changes implemented in aligning with APC and APNI.

## Sources

For each dataset in the compilation there is the option to list primary and secondary citations as well as, for compilations, original citations. The primary citation `r austraits$schema$metadata$elements$source$values$primary$description` while the secondary citation is `r austraits$schema$metadata$elements$source$values$secondary$description`. These references are included in two places:

1. Within the table [methods](#methods), where we provide a formatted version of each.
2. In the element [sources](#sources), where we provided bibtex versions of all sources which can be imported into your reference library. The keys for these references are listed within the [methods]{#methods}. 

# definitions {#traitdefs}

Following is a list of traits included in this version.

```{r, traits}
traits <- austraits$traits$trait_name %>% unique()
for(trait in traits) {
  elements <- austraits$definitions[[trait]]
  
  
  c( 
   sprintf("## %s\n", trait),
   sprintf("- label: %s", elements$label ),
   sprintf("- description: %s", elements$description ),
   sprintf("- type: %s%s", elements$type,  
      ifelse(elements$type == "numeric", sprintf("\n- units: %s", elements$units), "")), 
   ifelse(elements$type == "numeric", sprintf("- allowable range: %s - %s %s", elements$allowed_values_min, elements$allowed_values_max, elements$units), sprintf("- allowable values:\n%s\n",                paste0("    - *",elements$allowed_values_levels %>% names(), "*: ", elements$allowed_values_levels %>% unlist(), collapse="\n"))),
   "") %>% writeLines()
}
```
