---
title: "Dictionary for `AusTraits` data compilation - a curated plant trait database for the Australian flora"
output:
  html_document:
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: true
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

<!-- hack to get indentation on 3rd level of floating TOC; see
https://stackoverflow.com/questions/46201753/rmarkdown-indentation-of-toc-items-in-html-output
 -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# knitr defaults
knitr::opts_chunk$set(echo=FALSE, cache=FALSE, results='asis', message=FALSE, warning=FALSE)

# default for table format
options(knitr.table.format = "html")

# Guidelines for writing report code
# - use tidyverse style and format: http://htmlpreview.github.io/?https://github.com/nicercode/2018_BEES_regression/blob/master/tidyverse.html
# - use kableExtra for styling: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# - use knitr chunk options: https://rmarkdown.rstudio.com/lesson-3.html

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
library(tidyverse)
library(RefManageR)
library(knitr)
library(kableExtra)

my_kable_styling <- my_kable_styling_html

trait_definitions <- austraits$trait_definitions
schema <- austraits$schema
```

This document describes the structure of the AusTraits compilation, corresponding to **Version `r austraits$build_info$version`** of the dataset. 
Note that the information provided below is based on the information provided within the file `system.file("data", "schema.yml", package = "austraits.build")`. 

For details on access, structure and usage please visit https://doi.org/10.5281/zenodo.3568417

# Elements of AusTraits

The compiled AusTraits database has the following main components:

```
austraits
├── traits
├── sites
├── contexts
├── methods
├── excluded_data
├── taxonomic_updates
├── taxa
├── contributors
├── sources
├── trait_definitions
├── schema
└── build_info
```

These elements include all the data and contextual information submitted with each contributed datasets. Each component  is defined as follows:

```{r}
print_defintions_element <- function(elements) {
  if(elements$type == "character") {
    sprintf("**Content:** %s\n", elements$value) %>% 
      writeLines()
  }
  
  if(elements$type == "table") {
    
    sprintf("**Content:** \n") %>% 
      writeLines()

    elements$elements %>% 
      list1_to_df() %>%
      kableExtra::kable() %>% 
      writeLines()
  }
}

for(v in names(austraits)) {
  elements <- austraits$schema$austraits$elements[[v]]
  
  sprintf("## %s {#%s}\n\n**Description:** %s\n", v, v, elements$description) %>% 
    writeLines()
  
  elements %>%
    print_defintions_element()

  writeLines(c(""))
}
```


# Methological details {#methods}

## Dataset IDs

The core organising unit behind AusTraits is the `dataset_id`. Records are organisation as coming from a particular study, defined by the  `dataset_id`. Our preferred format for `dataset_id` is surname of the first author of any corresponding publication, followed by the year, as `surname_year`. E.g. `Falster_2005`. Wherever there are multiple studies with the same id, we add a suffix `_2`, `_3` etc. E.g.`Falster_2005`, `Falster_2005_2`.

## Observation IDs

As well as a `dataset_id`, each trait measurement has an associated `observation_id`.  Observation IDs bind together related measurements within any dataset, and thereby allow transformation between long  (e.g. with variables `trait_name` and `value`) and wide (e.g. with traits as columns) formats.

Generally, `observation_id` has the format `dataset_id_XX` where `XX` is a unique number within each dataset. For example, if multiple traits were collected on the same individual, the `observation_id` allows us to gather these together. For floras, which report a species averages, the `observation_id` is assigned at the species level.

For datasets that arrive in wide format we assume each row has a unique `observation_id`. For datasets that arrive in long format, the `observation_id` is assigned based on a specified grouping variable.  If missing, `observation_id` is assigned based on `taxon_name`. 

## Site names

As well as `dataset_id` and `observation_id`, where appropriate, trait values are associated with a `site_name`. Unique combinations of `dataset_id` and `site_name` can be used to cross-match against the sites table, which provide further details on the site sampled.

## Context names

As well as `dataset_id`, `observation_id`, and `site_name`, where appropriate, trait values are associated with a `context_name`. Unique combinations of `dataset_id` and `context_name` can be used to cross-match against the context table, which provide further details on the context sampled.

## Values and Value types {#value_types}

Each record in the table of trait data has an associated `value` and `value_type`. 

Traits are either `numeric` or `categorical`. For traits with numerical values, the recorded value has been converted into standardised units and we have check that the value can be converted into a number and lies within the allowable range.  For categorical variables, we only include records that are defined in the trait_definitions. Moreover, we use a format whereby

- we use `_` for multi-word terms, e.g. `semi_deciduous`
- use a space for situations where there are two possible values for that trait, e.g. `annual biennial` for something which is either annual or biennial 

Each trait measurement also an associated `value_type`, which gives ``r austraits$trait_definitions$value_type$description``. Possible values are:

```{r value_type}
austraits$trait_definitions$value_type$values %>% 
  list1_to_df()  %>% 
  my_kable_styling() %>% 
  writeLines()
```

AusTraits does not include intra-individual observations. When multiple measurements per individual are submitted to AusTraits, we take the mean of the values and record the value_type as `individual_mean`.

## Taxonomy

Version `r austraits$build_info$version` of AusTraits contains records for `r austraits$taxa %>% nrow()` different taxa. We have aligned taxa with known taxonomic units in the [`Australian Plant Census` (APC)](https://biodiversity.org.au/nsl/services/apc) and/or the [`Australian Plant Names Index` (APNI)](https://biodiversity.org.au/nsl/services/APNI). Of the `r austraits$taxa %>% nrow()` taxa included, `r austraits$taxa %>% filter(!is.na(source)) %>% nrow()` are aligned with known taxa. 

The `traits` table reports both the original and the updated taxon name alongside each trait record.

The table `taxa` lists all taxa in the database, including additional infomration about the taxa. 

The table `taxanomic_updates` provdies details on all taxonomic names changes implemented in aligning with APC and APNI.

## Sources

For each dataset in the compilation there is the option to list primary and secondary citations. The primary citation `r austraits$schema$metadata$elements$source$values$primary` while the secondary citation is `r austraits$schema$metadata$elements$source$values$secondary`. These references are included in two places:

1. Within the table [methods](#methods), where we provide a formatted version of each.
2. In the element [sources](#sources), where we provided bibtex versions of all sources which can be imported into your reference library. The keys for these references are listed within the [methods]{#methods}. 

# Trait_definitions {#traitdefs}

Followign is a list of traits included in this version.

```{r, traits}
traits <- austraits$traits$trait_name %>% unique()
for(trait in traits) {
  elements <- austraits$trait_definitions$traits$elements[[trait]]
  
  data_trait <- austraits$traits %>% filter(trait_name == trait)
  
  c( 
   sprintf("## %s\n", trait),
   sprintf("- label: %s", elements$label ),
   sprintf("- description: %s", elements$description ),
   sprintf("- type: %s%s", elements$type,  
      ifelse(elements$type == "numeric", sprintf("\n- units: %s", elements$units), "")), 
   ifelse(elements$type == "numeric", sprintf("- allowable range: %s - %s %s", elements$values$minimum,         elements$values$maximum, elements$units), sprintf("- allowable values:\n%s\n",                paste0("    - *",elements$values %>% names(), "*: ", elements$values %>% unlist(), collapse="\n"))),
   sprintf("- number of records: %s", data_trait %>% nrow() ),
   sprintf("- number of studies: %s", data_trait %>% pull(dataset_id) %>% unique() %>% length() ),
   "") %>% writeLines()
}
```
