---
params:
  v_prev: NA
  v_curr: NA
output: github_document
keep_md: TRUE
---

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_root("NEWS.md"))

library(tidyverse)
library(RefManageR)

clean <- function(v) gsub("v", "", v, fixed=TRUE)
vclean <- function(v) paste0("v", clean(v))

v_prev <- clean(params$v_prev)
v_curr <- clean(params$v_curr)
# for debug pruposes
# params <- list()
# params$v_prev <- "3.0.2"
# params$v_curr <- "3.0.2.9000"

# Create news
filename <- function(v) {
  here::here(sprintf("export/data/austraits-%s/austraits-%s.rds", v, v))
}

austraits_prev <- readRDS(filename(v_prev))
austraits_curr <- readRDS(filename(v_curr))

# patches to make various versions play nicely together

if(v_prev <= "3.0.2") {
    austraits_prev$traits[["location_id"]] <- austraits_prev$traits[["site_name"]]
}
austraits_curr$contributors <- austraits_curr$contributors %>% mutate(name = paste(given_name, last_name))

```


# AusTraits v`r v_curr`

This is major/minor release of the AusTraits database. 

- `r sprintf("austraits-%s.zip", v_curr)`: contains the compiled dataset and detailed of structure
- `r sprintf("austraits-%s.rds", v_curr)`: contains a version of the dataset for direct loading in R
- source code `r sprintf("v%s.zip", v_curr)`: contains the source materials used to build the compiled dataset

For details on access, structure and usage please visit <https://doi.org/10.5281/zenodo.3568417>

This release was generated from source materials available at `r sprintf("https://github.com/traitecoevo/austraits.build/releases/tag/v%s", v_curr)`
A full set of changes in the source can be viewed at: `r sprintf("https://github.com/traitecoevo/austraits.build/compare/v%s...v%s", v_prev, v_curr)` 

Compared to the last version, this release contains substantial additions of new data and improvement of old data. 

```{r, results="asis"}

f_summarise <- function(data, v) { 
  data$traits %>% 
  summarise(version = data$build_info$version %>% as.character,
            dataset_id = n_distinct(dataset_id),
            taxa = n_distinct(taxon_name),
            locations = n_distinct(location_id),
            traits = n_distinct(trait_name),
            records= n()
  )
}

bind_rows(
  austraits_prev %>% f_summarise(v_prev),
  austraits_curr %>% f_summarise(v_curr)
) %>% knitr::kable("markdown") %>% writeLines()
```


This release contains:

```{r, results='asis'}
remove_studies <- austraits_prev$traits %>% filter(!(dataset_id %in% austraits_curr$traits$dataset_id)) %>% pull(dataset_id) %>% unique()
new_studies <- austraits_curr$traits %>% filter(!(dataset_id %in% austraits_prev$traits$dataset_id)) %>% pull(dataset_id) %>% unique()

new_data <- austraits_curr$traits %>% filter(!(dataset_id %in% austraits_prev$traits$dataset_id))
new_traits <- new_data %>% filter(!trait_name %in% austraits_prev$traits$trait_name) %>% pull(trait_name) %>% unique() %>% sort()

new_people <- austraits_curr$contributors %>% filter(!name %in% austraits_prev$contributors$name) %>% pull(name) %>% unique() %>% sort() 
```


- **`r length(new_studies)` new sources added to the dataset**: `r new_studies %>% paste(collapse = ", " )`
- **`r length(new_people)` new contributors:** `r new_people %>% paste(collapse = ", " )`

- **`r length(remove_studies)` sources removed**-- most replaced by corrected 
  or updated sources above: `r remove_studies %>% paste(collapse = ", " )`

- **`r new_data %>% filter(trait_name %in% austraits_prev$traits$trait_name) %>% nrow()` new or updated records for `r new_data %>% filter(trait_name %in% austraits_prev$traits$trait_name) %>% pull(trait_name) %>% n_distinct()` existing traits**

- **`r new_data %>% filter(!trait_name %in% austraits_prev$traits$trait_name) %>% nrow()` new records for `r new_traits %>% n_distinct()` new traits:** `r new_traits %>% paste(collapse = ", " )`


Details on new studies: 

```{r, results="asis"}
new_data %>% 
	group_by(dataset_id) %>% 
	summarise(taxa = length(unique(taxon_name)), 
	          locations = length(unique(location_id)),
	          traits = length(unique(trait_name)),
	          records= n()
	          ) %>% knitr::kable("markdown") %>% writeLines()
```

