---
title: Working with the `AusTraits` GitHub repo
author: Daniel Falster
date: "`r Sys.Date()`"
header-includes:
   - \usepackage{pmboxdraw}
output:
  github_document:
    toc: yes
    toc_depth: 4
  html_document:
    keep_md: yes
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: console
---

<!-- Working_with_github.md is generated from Working_with_github.Rmd Please edit that file -->

<!-- hack to get indentation on 3rd level of floating TOC in html; see
https://stackoverflow.com/questions/46201753/rmarkdown-indentation-of-toc-items-in-html-output
 -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# knitr defaults
root.dir = rprojroot::find_root("remake.yml")
knitr::opts_knit$set(root.dir = root.dir)

knitr::opts_chunk$set(
  echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE,
  results='asis',
  fig.cap='', 
  fig.path = file.path(root.dir,"docs/figures", paste0("working","_")),
  out.width = "100%"
)
```


This document outlines our processes for handling various activities in GitHub.

# Preparing a new dataset



# Commit messages

Informative commit messages are ideal. Where possible, these should link to 



# Pull requests

To makes changes you'll need to send a pull request (PR) on GitHub. For an intro to PRs [see here](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/about-pull-requests). In short, 

1. Create a Git branch for your new work, either within the AusTraits repo (if you are an approved contributor) or as a [fork of the repo](https://help.github.com/en/github/getting-started-with-github/fork-a-repo). 
2. Make commits and push these up onto the branch. 
2. Make sure everything runs fine before you send a pull request.
3. When you're ready to merge in the new features, 

Before you make a substantial pull request, you should always [file an issue](https://github.com/traitecoevo/austraits.build/issues) and make sure someone from the team agrees that it’s worth pursuing. a problem. If you’ve found a bug, create an associated issue and illustrate the bug with a minimal [reprex](https://www.tidyverse.org/help/#reprex) illustrating the issue.


## Merging a pull request

There are multiple ways to merge a pull request, including using GitHub's built-in options for merging and squashing. When merging a PR, we ideally want

- a single commit
- attributing the work to the original author
- to run various checks along the way 

There's two ways to do this. For both you need to be an approved maintainer. 

### Merging in your own PR

You can merge in your own PR after you've had someone else review it. 

1. Send the PR
2. Tag someone to review
3. Once ready, merge into main choosing "Squash & Merge", using an informative commit message.


### Merging someone else's PR

When merging in someone else's PR, the buitlin  options aren't ideal. as they either take all of the commits on a branch (ugh, messy), OR make the commit under the name of the person merging the request.

The workflow below describes how to merge a pull request from the command line, with a single commit & attributing the work to the original author. Lets assume a branch of name `Smith_1995`.

First from the master branch in the repo, run the following:

```
git merge --squash origin/Smith_1995
```

Then in R

```{r, eval=FALSE}
dataset_ids <- "Smith_1995"

# Update remake file
austraits_rebuild_remake_setup()

# check data builds
remake::make(dataset_ids)

# Check taxonomy has been updated
metadata_check_taxa(dataset_ids)

# run tests on dataset
austraits_run_tests()

# read and write yaml to prevent future reformatting 
f <- file.path("data", dataset_ids, "metadata.yml")
read_yaml(f) %>% write_yaml(f)

# rebuild
remake::make(dataset_ids)
```

Now back in the terminal

```
git add .
git commit
```

Add a commit message, referencing relevant pull requests and issues, e.g.

```
Smith_1995: Import new data

For #224, closes #286
```

And finally, amend the commit author, to reference the person who did all the work!
```
git commit --amend --author "XXX <XXX@gmail.com>"
```

# Making a new release & version updating

Notes on release process

http://r-pkgs.had.co.nz/release.html

## Releasing a version

Increment version number

```{r, eval=FALSE}
desc::desc_bump_version("dev")
```

XXXXX
